---
title: "Machine Learning Coursera"
output: html_document
---


##Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 



##Data 


The training data for this project are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment. 

##What you should submit

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases. 

1. Your submission should consist of a link to a Github repo with your R markdown and compiled HTML file describing your analysis. Please constrain the text of the writeup to < 2000 words and the number of figures to be less than 5. It will make it easier for the graders if you submit a repo with a gh-pages branch so the HTML page can be viewed online (and you always want to make it easy on graders :-).
2. You should also apply your machine learning algorithm to the 20 test cases available in the test data above. Please submit your predictions in appropriate format to the programming assignment for automated grading. See the programming assignment for additional details. 


##Reprduceability
The seed was set to 2345 as seen in the code. All the packages that were used are going to be seen in the code as well.

##Cross Validation
I will divide the training data to 2 sets 70% will be for the training and 30% will be for testing each model. 

##Out of Sample Error
The expected out-of-sample error will correspond to the quantity: 1-accuracy in the cross-validation data. Accuracy is the proportion of correct classified observation over the total sample in the subTesting data set. Expected accuracy is the expected accuracy in the out-of-sample data set (i.e. original testing data set). Thus, the expected value of the out-of-sample error will correspond to the expected number of missclassified observations/total observations in the Test data set, which is the quantity: 1-accuracy found from the cross-validation data set.

#Code
Loading Packages and setting the seed
```{r}
library(caret)
library(randomForest)
library(MASS)
library(rpart)
library(rpart.plot)
library(rattle)
set.seed(2345)
```

Reading the files


```{r}
training<-read.csv("train.csv",na.strings = c("",NA))
testing<-read.csv("test.csv",na.strings = c("",NA))
```


Removing unwanted columns (NAs and unnecessary)

```{r,results='hide'}
training<-training[,colSums(is.na(training))==0]
testing<-testing[,colSums(is.na(testing))==0]
training<-training[,-c(1:7)]
testing<-testing[,-c(1:7)]
dim(training)
dim(testing)
```

Cross Validation
```{r}
intrain<-createDataPartition(y=training$classe,p=0.7,list=FALSE)
training.train<-training[intrain,]
training.test<-training[-intrain,]
dim(training.train)
dim(training.test)
```

##First Model
```{r}
model1<-train(classe~.,data = training.train,method="lda")
predict1<-predict(model1,training.test)
confusionMatrix(predict1,training.test$classe)
```
##Second Model
```{r,cache=TRUE}
model2<-randomForest(classe~.,data = training.train)
predict2<-predict(model2,training.test)
confusionMatrix(predict2,training.test$classe)
```
##Third Model
```{r,cache=TRUE}
model3<-train(classe~.,data = training.train,method="rpart")
predict3<-predict(model3,training.test)
confusionMatrix(predict3,training.test$classe)
fancyRpartPlot(model3$finalModel)
```
##Decision
We can see we got the most accurate results, and by far, when using model 2, the random forest model. We got an accuracy of 0.9933 as opposed to 0.7 in the first model and 0.4984 in the third.
The expected out of sample error is 1-Accuracy = 0.0067. With such high accuracy we will expect to predict correctly most if not all of the test set.

##Predicted Results
```{r}
prediction<-predict(model2,testing)
prediction
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
  pml_write_files(prediction)
}
```